---
title: ""
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{breakout}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-------------------------->
<!-------------------------->
# Setup
<!-------------------------->
<!-------------------------->
Before running rgcambreakout you will need:

* A running version of GCAM
* IEA_EnergyBalances_2019.csv file placed in the ./input/inst/extdata/energy
* CEDS data places in input/gcamdata/inst/extdata/emissions/CEDS

<!-------------------------->
<!-------------------------->
# Breakout a new region
<!-------------------------->
<!-------------------------->

This script lets users breakout a new region which can be a combination of or a single country from the list of countries found in the file: "/inst/extdata/common/iso_GCAM_regID.csv" under the column country_name. The following example shows how to breakout 'Thailand' and 'Lao Peoples Democratic Republic' as a new region called 'Thailand Laos'.

```{r, eval=F}
library(rgcambreakout)
breakout(gcamdataFolder = "PATH_TO_GCAM_FOLDER/input/gcamdata",
         regionNew = "Thailand Laos",
         countriesNew = c("Thailand","Lao Peoples Democratic Republic"))
```

After the region is broken out the following files will have changed. A copy of each of these files appended with '_Original' will remain in the same folders:

* /inst/extdata/common/iso_GCAM_regID.csv
* /inst/extdata/common/GCAM_region_names.csv
* /inst/extdata/aglu/A_bio_frac_prod_R.csv
* /inst/extdata/aglu/A_soil_time_scale_R.csv
* /inst/extdata/emissions/A_regions.csv
* /inst/extdata/energy/A23.subsector_interp_R.csv
* /inst/extdata/energy/A_regions.csv
* /inst/extdata/energy/offshore_wind_potential_scaler.csv

**After running this script you will need to re-build the gcamdata system by "install & restart" from R Studio followed by driver().**

<!-------------------------->
<!-------------------------->
# Restore Original Files
<!-------------------------->
<!-------------------------->

Users can also restore back to the original files by running the following code. This will replace the newly created files with the versions appended with '_Original' and delete the appended files. The gcamdata system will have to be rebuilt using 'install & restart' from R Studio once again followed by driver(). 

```{r, eval=F}
library(rgcambreakout)
restore(gcamdataFolder = "PATH_TO_GCAM_FOLDER/input/gcamdata")
```
