---
title: "Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

<!-------------------------->
<!-------------------------->
# Setup
<!-------------------------->
<!-------------------------->
Before running gcambreakout you will need:

* A running version of GCAM
* IEA_EnergyBalances_2019.csv file placed in the ./input/inst/extdata/energy
* CEDS data placed in input/gcamdata/inst/extdata/emissions/CEDS
* NOTE: The CEDS data and IEA_EnergyBalances_2019.csv is proprietary data and only accesible to internal JGCRI staff at https://stash.pnnl.gov/scm/jgcri/gcam-proprietary-data.git.

The breakout function can be used to breakout new regions with countries and further subRegions as follows:

* `regionNew`: Region New will breakout a larger region which can be a combination of available countries in  "/inst/extdata/common/iso_GCAM_regID.csv" and which must be listed in param `countriesNew`
* `countriesNew`: Is the list of countries which makes up the `regionNew`
* `breakoutCountriesNew`: Is set to `F` as the default. When turned on (`T`) all `countriesNew` within `regionNew` will also be broken out with their own socio-economics, electricity, buildings, transport data as well as any trade between them if it exists. Individual sectors can be turned off for this country level disaggregation if desired as follows.
* `breakoutCountriesNew_elec`: Default is `T`. If set to `F` then electricity will not be broken out for the countries within the `regionNew`.

<!-------------------------->
<!-------------------------->
# Breakout a new multi-country region
<!-------------------------->
<!-------------------------->

This script lets users breakout a new region which can be a combination of or a single country from the list of countries found in the file: "/inst/extdata/common/iso_GCAM_regID.csv" under the column country_name. The following example shows how to breakout 'Thailand' and 'Lao Peoples Democratic Republic' as a new region called 'Thailand Laos'.

```{r, eval=F}
library(gcambreakout)
breakout(gcamdataFolder = "PATH_TO_GCAM_FOLDER/input/gcamdata",
         regionNew = "Thailand Laos",
         countriesNew = c("Thailand","Lao Peoples Democratic Republic"))
```

After the region is broken out the following files will have changed. A copy of each of these files appended with '_Original' will remain in the same folders:

* /inst/extdata/common/iso_GCAM_regID.csv
* /inst/extdata/common/GCAM_region_names.csv
* /inst/extdata/aglu/A_bio_frac_prod_R.csv
* /inst/extdata/aglu/A_soil_time_scale_R.csv
* /inst/extdata/emissions/A_regions.csv
* /inst/extdata/energy/A23.subsector_interp_R.csv
* /inst/extdata/energy/A_regions.csv
* /inst/extdata/energy/offshore_wind_potential_scaler.csv

**After running this script you will need to re-build the gcamdata system by "install & restart" from R Studio followed by driver().**

**NOTE:** 

* **The countries or subregions within the new regions will not have individual data in the model outputs. Only the combined new region will report outputs.**
**NOTE: The regionNew will not include electricity trade with other regions**


<!-------------------------->
<!-------------------------->
# Breakout countries within multi-country region
<!-------------------------->
<!-------------------------->

When `breakoutCountriesNew` is set to `T` (Default is `F`) this functionality will be activated and each individual `countriesNew` within the larger `regionNew` will be broken out. This is useful when users want to represent the individual `countriesNew` within the larger `regionNew` as well as for capturing dynamics between the `countriesNew`. The following files will change:

* Socio-economics
* Electricity sector
  * S123: processing of electricity sector inputs and outputs by fuel type. includes cogen (all zeroes)
  * S1231: disaggregation of ng single cycle from ng combined cycle
  * S2233: generation of tables for XML. Use relative shares from S1231 to downscale L2233 tables.
  * should be able to shortcut L223 electricity; going instead from L2233.electricity_water.
* End use sectors - general
  * S131: general end use partitioning. processing should generally follow LA131.
  * electricity demands calculated as:
    * generation minus ownuse and T&D losses,
    * minus unscalable demands (e.g., refining sector),
    * multiplied by each sector's (bld, ind, trn) share of total elec demands for the combined region
* Industry
  * S132: industrial sector demands (energy downscaled from S131, intensities passed down)
  * S232: copy structure of LA132, re-set electricity market to combined region
* Buildings
  * S144: buildings sector (energy downscaled from L131, intensities passed down)
  * S244: copy structure of L244, re-set electricity market to combined region
* Transportation
  * S154: transportation sector (energy downscaled from L131, intensities passed down)
  * S254: copy structure of L254, re-set electricity market to combined region

<!-------------------------->
<!-------------------------->
# Restore Original Files
<!-------------------------->
<!-------------------------->

Users can also restore back to the original files by running the following code. This will replace the newly created files with the versions appended with '_Original' and delete the appended files. The gcamdata system will have to be rebuilt using 'install & restart' from R Studio once again followed by driver(). 

```{r, eval=F}
library(gcambreakout)
restore(gcamdataFolder = "PATH_TO_GCAM_FOLDER/input/gcamdata")
```


