---
title: "Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

<!-------------------------->
<!-------------------------->
# Setup
<!-------------------------->
<!-------------------------->
Before running gcambreakout you will need:

* A running version of GCAM
* IEA_EnergyBalances_2019.csv file placed in the ./input/inst/extdata/energy
* CEDS data placed in input/gcamdata/inst/extdata/emissions/CEDS
* NOTE: The CEDS data and IEA_EnergyBalances_2019.csv is proprietary data and only accesible to internal JGCRI staff at https://stash.pnnl.gov/scm/jgcri/gcam-proprietary-data.git.

The breakout function can be used to breakout new regions with countries and further subRegions as follows:

* `regionNew`: Region New will breakout a larger region which can be a combination of available countries in  "/inst/extdata/common/iso_GCAM_regID.csv" and which must be listed in param `countriesNew`
* `countriesNew`: Is the list of countries which makes up the `regionNew`
* `breakoutCountriesNew`: Is set to `F` as the default. When turned on (`T`) all `countriesNew` within `regionNew` will also be broken out with their own socio-economics, electricity, buildings, transport data as well as any trade between them if it exists. Individual sectors can be turned off for this country level disaggregation if desired as follows.
* `breakoutCountriesNew_elec`: Default is `T`. If set to `F` then electricity will not be broken out for the countries within the `regionNew`.


<!-------------------------->
<!-------------------------->
# Restore Original Files
<!-------------------------->
<!-------------------------->

Users can also restore back to the original files by running the following code. This will replace the newly created files with the versions appended with '_Original' and delete the appended files. The gcamdata system will have to be rebuilt using 'install & restart' from R Studio once again followed by driver(). 

```{r, eval=F}
library(gcambreakout)
restore(gcamdataFolder = "PATH_TO_GCAM_FOLDER/input/gcamdata")
```


