---
title: "breakout"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{breakout}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-------------------------->
<!-------------------------->
# Key Links
<!-------------------------->
<!-------------------------->

- Github: https://github.com/JGCRI/rgcambreakout
- Webpage: https://jgcri.github.io/rgcambreakout/
- GCAM Adding a new region: https://github.com/JGCRI/gcamdata/wiki/Modifying-GCAM-via-the-Data-System
- GCAM proprietary data (Will need access): https://stash.pnnl.gov/projects/JGCRI/repos/gcam-proprietary-data/browse


<!-------------------------->
<!-------------------------->
# Setup
<!-------------------------->
<!-------------------------->
Before running rgcambreakout you will need:
- A running version of GCAM
- IEA_EnergyBalances_2019.csv file placed in the ./input/inst/extdata/energy
- CEDS data places in input/gcamdata/inst/extdata/emissions/CEDS

<!-------------------------->
<!-------------------------->
# Breakout a new region
<!-------------------------->
<!-------------------------->

This script lets users breakout a new region which can be a combination of or a single country from the list of countries found in the file: "/inst/extdata/common/iso_GCAM_regID.csv" under the column country_name. The list includes: 

"Aruba, Afghanistan, Angola, Anguilla, Albania, Andorra, Netherlands Antilles, United Arab Emirates, Argentina, Armenia, American Samoa, Antigua & Barbuda, Australia, Austria, Azerbaijan, Burundi, Belgium, Benin, Burkina Faso, Bangladesh, Bulgaria, Bahrain, Bahamas, Bosnia and Herzegovina, Belarus, Belize, Bermuda, Bolivia, Brazil, Barbados, Brunei Darussalam, Bhutan, Botswana, Central African Republic, Canada, Cocos (Keeling) Islands, Switzerland, Channel Islands, Chile, China, Cote dIvoire, Cameroon, Congo, the Democratic Republic of the, Congo, Cook Islands, Colombia, Comoros, Cape Verde, Costa Rica, Cuba, Curacao, Cayman Islands, Cyprus, Christmas Island, Czech Republic, Germany, Djibouti, Dominica, Denmark, Dominican Republic, Algeria, Ecuador, Egypt, Eritrea, Western Sahara, Spain, Estonia, Ethiopia, Finland, Fiji, Falkland Islands (Malvinas), France, Faroe Islands, Micronesia, Federated States of, Gabon, United Kingdom, Georgia, Ghana, Gibraltar, Guinea, Guadeloupe, Gambia, Guinea-Bissau, Equatorial Guinea, Greece, Grenada, Greenland, Guatemala, French Guiana, Guam, Guyana, Hong Kong, Honduras, Croatia, Haiti, Hungary, Indonesia, Isle of Man, India, Ireland, Iran, Islamic Republic of, Iraq, Iceland, Israel, Italy, Jamaica, Jordan, Japan, Kazakhstan, Kenya, Kyrgyzstan, Cambodia, Kiribati, Saint Kitts and Nevis, Korea, Republic of, Kuwait, Lao Peoples Democratic Republic, Lebanon, Liberia, Libyan Arab Jamahiriya, Saint Lucia, Liechtenstein, Sri Lanka, Lesotho, Lithuania, Luxembourg, Latvia, Macau, Morocco, Monaco, Moldova, Republic of, Madagascar, Maldives, Mexico, Marshall Islands, Macedonia, the former Yugoslav Republic of, Mali, Malta, Myanmar, Montenegro, Mongolia, Northern Mariana Islands, Mozambique, Mauritania, Montserrat, Martinique, Mauritius, Malawi, Malaysia, Mayotte, Namibia, New Caledonia, Niger, Norfolk Island, Nigeria, Nicaragua, Niue, Netherlands, Norway, Nepal, Nauru, New Zealand, Oman, Pakistan, Panama, Pacific Islands Trust Territory, Pitcairn Islands, Peru, Philippines, Palau, Papua New Guinea, Poland, Puerto Rico, Korea, Democratic Peoples Republic of, Portugal, Paraguay, Palestinian Territory, Occupied, French Polynesia, Qatar, Reunion, Romania, Russian Federation, Rwanda, Saudi Arabia, Serbia and Montenegro, Sudan, Senegal, Singapore, Saint Helena, Svalbard and Jan Mayen Islands, Solomon Islands, Sierra Leone, El Salvador, San Marino, Somalia, Saint Pierre and Miquelon, Serbia, South Sudan, Sao Tome and Principe, Suriname, Slovakia, Slovenia, Sweden, Swaziland, Seychelles, Syrian Arab Republic, Turks and Caicos, Chad, Togo, Thailand, Tajikistan, Tokelau, Turkmenistan, Timor Leste, Tonga, Trinidad and Tobago, Tunisia, Turkey, Tuvalu, Taiwan, Tanzania, United Republic of, Uganda, Ukraine, Uruguay, United States of America, Uzbekistan, Vatican, Saint Vincent and the Grenadines, Venezuela, Virgin Islands, British, Virgin Islands, U.S., Viet Nam, Vanuatu, Wallis and Futuna, Samoa, Yemen, Yugoslavia, Federal Republic of, South Africa, Zambia, Zimbabwe"

The following example shows how to breakout 'Thailand' and 'Lao Peoples Democratic Republic' as a new region called 'Thailand Laos'.

```r
breakout(gcamdataFolder = "PATH_TO_GCAM_FOLDER/input/gcamdata",
         regionNew = "Thailand Laos",
         countriesNew = c("Thailand","Lao Peoples Democratic Republic"))
```

After the region is broken out the following files will have changed. A copy of each of these files appended with '_Original' will remain in the same folders:

* /inst/extdata/common/iso_GCAM_regID.csv
* /inst/extdata/common/GCAM_region_names.csv
* /inst/extdata/aglu/A_bio_frac_prod_R.csv
* /inst/extdata/aglu/A_soil_time_scale_R.csv
* /inst/extdata/emissions/A_regions.csv
* /inst/extdata/energy/A23.subsector_interp_R.csv
* /inst/extdata/energy/A_regions.csv
* /inst/extdata/energy/offshore_wind_potential_scaler.csv

**After running this script you will need to rebuild the gcamdata system by "install & restart" from R Studio followed by driver().**

Users can also restore back to the original files by running the following code. This will replace the newly created files with the versions appended with '_Original'. The gcamdata system will have to be rebuilt using 'install & restart' from R Studio once again followed by driver(). 

```r
restore(gcamdataFolder = "PATH_TO_GCAM_FOLDER/input/gcamdata")
```


```{r setup}
library(rgcambreakout)
```
